---
title: "Progress Memo 2"
subtitle: |
  | Final Project 
  | Data Science 1 with R (STAT 301-1)
author: "Celena Kim"
date: today

format:
  html:
    toc: true
    embed-resources: true
    code-fold: show
    link-external-newwindow: true
    
execute:
  warning: false

from: markdown+emoji 
---


::: {.callout-tip icon="false"}
## Github Repo Link

[https://github.com/stat301-1-2023-fall/final-project-1-celenakim](https://github.com/stat301-1-2023-fall/final-project-1-celenakim)
:::

## Overview

Students are expected to demonstrate that significant progress has been made on their final project since the submission of progress memo 1. Students should have their data cleaned and the EDA should be started.

Demonstrating significant progress means students should have some univariate and bivariate analyses complete for several of their variables.  They should share a few graphics and/or tables with a description of what they have found thus far to demonstrate they progress. Students should should clearly state what they are exploring and why in these demonstrations. That is, they should share the guiding curiosity or research question that accompanies the particular graphics and/or tables they choose to share. 

What else should be in the memo
Students should summarize their progress, where they are at, and what their next steps will be. Self assessment of progress would also be appropriate. When thinking or describing next steps students should share any guiding curiosities or research questions they plan to explore.

## Packages & Datasets
The data used in this project is called [Hollywood Hits and Flops (2007 - 2023)](https://www.kaggle.com/datasets/sujaykapadnis/hollywood-hits-and-flops-2007-2023/) found on [Kaggle](https://www.kaggle.com). 

```{r}
#| label: load-pkgs-data

# Loading packages and data
library(tidyverse)
data_2007 <- read.csv("data/The Hollywood Inider - all data - 2007.csv")
data_2008 <- read.csv("data/The Hollywood Inider - all data - 2008.csv")
data_2009 <- read.csv("data/The Hollywood Inider - all data - 2009.csv")
data_2010 <- read.csv("data/The Hollywood Inider - all data - 2010.csv")
data_2011 <- read.csv("data/The Hollywood Inider - all data - 2011.csv")
data_2012 <- read.csv("data/The Hollywood Inider - all data - 2012.csv")
data_2013 <- read.csv("data/The Hollywood Inider - all data - 2013.csv")
data_2014 <- read.csv("data/The Hollywood Inider - all data - 2014.csv")
data_2015 <- read.csv("data/The Hollywood Inider - all data - 2015.csv")
data_2016 <- read.csv("data/The Hollywood Inider - all data - 2016.csv")
data_2017 <- read.csv("data/The Hollywood Inider - all data - 2017.csv")
data_2018 <- read.csv("data/The Hollywood Inider - all data - 2018.csv")
data_2019 <- read.csv("data/The Hollywood Inider - all data - 2019.csv")
data_2020 <- read.csv("data/The Hollywood Inider - all data - 2020.csv")
data_2021 <- read.csv("data/The Hollywood Inider - all data - 2021.csv")
data_2022 <- read.csv("data/The Hollywood Inider - all data - 2022.csv")
```

## Combining Data
```{r}
# data_2019 <- data_2019 |> 
#   rename(Genre = Genres, Opening.Weekend = Opening.Weekend...., of.Gross.earned.abroad = X..of.Gross.earned.abroad)
# data_2019 <- data_2019[, -ncol(data_2019)]
# 
# data_2020 <- data_2020 |> 
#   rename(Genre = Genres, Opening.Weekend = Opening.Weekend...., of.Gross.earned.abroad = X..of.Gross.earned.abroad) 
# data_2020 <- data_2020[, -ncol(data_2020)]
# 
# data_2021 <- data_2021 |> 
#   rename(Genre = Genres, Opening.Weekend = Opening.Weekend...., of.Gross.earned.abroad = X..of.Gross.earned.abroad) 
# data_2021 <- data_2021[, -ncol(data_2021)]
# 
# data_2022 <- data_2022 |> 
#   rename(Genre = Genres, Opening.Weekend = Opening.Weekend...., of.Gross.earned.abroad = X..of.Gross.earned.abroad) 
# data_2022 <- data_2022[, -ncol(data_2022)]
# 
# movie_data <- rbind(data_2011, data_2012, data_2013, data_2014, data_2015, data_2016, data_2017, data_2018, data_2019, data_2020, data_2021, data_2022) #2011-2022
# 
# write.csv(movie_data, "movie_data.csv", row.names = FALSE)
# movie_data <- movie_data[, -ncol(movie_data)]

```

```{r}
data_2011 <- data_2011[, -ncol(data_2011)]

data_2012 <- data_2012[, -ncol(data_2012)]

data_2013 <- data_2013[, -ncol(data_2013)]

data_2014 <- data_2014[, -ncol(data_2014)]

data_2015 <- data_2015[, -ncol(data_2015)]

data_2016 <- data_2016[, -ncol(data_2016)]

data_2017 <- data_2017[, -ncol(data_2017)]

data_2018 <- data_2018[, -ncol(data_2018)]

data_2019 <- data_2019 |> 
  rename(Genre = Genres, Opening.Weekend = Opening.Weekend...., of.Gross.earned.abroad = X..of.Gross.earned.abroad)
data_2019 <- data_2019[, -c((ncol(data_2019) - 1):ncol(data_2019))]

data_2020 <- data_2020 |> 
  rename(Genre = Genres, Opening.Weekend = Opening.Weekend...., of.Gross.earned.abroad = X..of.Gross.earned.abroad) 
data_2020 <- data_2020[, -c((ncol(data_2020) - 1):ncol(data_2020))]

data_2021 <- data_2021 |> 
  rename(Genre = Genres, Opening.Weekend = Opening.Weekend...., of.Gross.earned.abroad = X..of.Gross.earned.abroad) 
data_2021 <- data_2021[, -c((ncol(data_2021) - 1):ncol(data_2021))]

data_2022 <- data_2022 |> 
  rename(Genre = Genres, Opening.Weekend = Opening.Weekend...., of.Gross.earned.abroad = X..of.Gross.earned.abroad) 
data_2022 <- data_2022[, -c((ncol(data_2022) - 1):ncol(data_2022))]

movie_data <- rbind(data_2007, data_2008, data_2009, data_2010, data_2011, data_2012, data_2013, data_2014, data_2015, data_2016, data_2017, data_2018, data_2019, data_2020, data_2021, data_2022) #2011-2022

write.csv(movie_data, "movie_data.csv", row.names = FALSE)
```

primary genre and genre NA
